\subsection{Related Work}
\label{sec:related}

To our knowledge, MiGen's TA tools represent the first work ---
preliminary results were published in~\cite{TA-ECTEL} --- that is
targeted at notifying teachers of students’ progress and state during
exploratory learning activities in the classroom, notifying the
teacher of students’ attainment of key indicators, and aiming to
inform the teacher’s own interventions in the class. This novelty of
our TA tools has presented a number of methodological challenges,
which we discuss in Section 3. In the last two years, several similar
initiatives have appeared, including an approach
recently shared by~\cite{Gutierrez12} ---inspired by early work
by~\cite{Yardi08}--- that focuses on progress measurement and
whether students are stuck in the context of programming labs.
Earlier work (e.g.~\cite{Gueraud09}) focuses mostly on the statistics
 of the interaction (e.g.~how often did the student produce a certain
 kind of indicator). As explained in this paper, the requirement
 analysis with teachers showed that immediate on-line feedback
 is more valuable than simple statistics as it allows orchestrating
 the use of technologies in the classroom.\ednote{MM: I don't understnad
   the comparison with the statistics --- maybe because I haven't read
   the work --- also not sure we ever say again what we mean with
   on-line feedback. Do you mean that the work of Gueraud09 is
   off-line? SG: Yes.}

The trend towards teacher support appears to grow
also in the learning analytics community
(see for example~\cite{Crespo12,Zaldivar12,Pardo12}, and there is high
synergetic potential between their work and the work reported here.
Other related initiatives include 
using Web log data generated by course management systems
(e.g.~WebCT) to help instructors become aware of students’
activities in distance learning classes~\cite{Mazza07}; post-analysis of
the system's data logs for helping teachers understand students'
behaviour in adaptive tutorials~\cite{BenAnim08}; 
 providing
awareness information to teachers so as to support their role as 
moderators of multiple e-discussions~\cite{Wichmann09} or class-wide
collaborative activities supported by hand-held devices~\cite{CortezNussbaum2009}. 
%
Particularly interesting is the work described in 
is~\cite{Avouris08}, that uses tools to analyse CSCL 
synchronous interaction to help the teacher; their use of ``rules'' to
find specific landmarks in the interaction bears some
similarity to the search of indicators we report here. 
%
However, none of these works
focuses on exploratory learning activities specifically. 

% Stefan Weinbrenner, Jan Engler, Astrid Wichmann, Ulrich Hoppe
% Monitoring and Analyzing Students' Systematic Behaviour by Agents -
% The SCY Pedagogical Agent Framework. Proceedings of the 5th European
% Conference on Technology Enhanced Learning (EC-TEL 2010), Barcelona,
% ES, 2010

% -- this I think is one of those mentioned at the top
% Stefan Weinbrenner, Jan Engler, Astrid Wichmann, Ulrich Hoppe
% Monitoring and Analyzing Students' Systematic Behaviour by Agents -
% The SCY Pedagogical Agent Framework. Proceedings of the 5th European
% Conference on Technology Enhanced Learning (EC-TEL 2010), Barcelona,
% ES, 2010


% Discussion here also of other related work on: 
% \begin{itemize}
% \item  interaction analysis
% \item  work on CSCL (even if we are not targetting collaboration here)
%   to support teachers 
% \item  work on teachers’ tools for moderating online discussions and
%   argumentation 
% \item  other tools for learning
% \item Argonaut's teacher tools (Astrid Wichman, Ulrich Hoppe)
% \end{itemize}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"